---
title: "ALDEX2 + PCA plots and tree of assembed mcr contigs"
output:
  html_document:
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(compositions)
library(ALDEx2)
library(stringr)
library(plyr)
library(ggnewscale)
library(ggalt)
library(RColorBrewer)
library(ggrepel)
library(tidyverse)
library(ape)
library(igraph)
library(phytools)
library(ggtree)
library(gggenes)
library(ggnewscale)
library(gtable)
library(grid)
library(kableExtra)
library(ggtreeExtra)
library(ggnewscale)
library(ggpubr)
library(genbankr)
library(cowplot)

countryRegions <- rworldmap::countryRegions


path='~/repos/mcr_metagenomes/'
knitr::opts_knit$set(root.dir = normalizePath(path)) 

source(paste(path, 'src/extra_funcs.R', sep='/'))
source(paste(path, 'src/ttests.R', sep='/'))
source(paste(path, 'src/pca_plots.R', sep='/'))
```

# ALDex2 + PCA plots
## Load data
```{r}
file <- "data/mcr_pivot_runs.csv"
mcr.data <- read.csv(file, check.names = F)
row.names(mcr.data) <- mcr.data$run_accession
mcr.data$collection_date <- as.Date(mcr.data$collection_date, '%Y-%m-%d')
mcr.data$collection_year <- format(as.Date(mcr.data$collection_date, '%Y-%m-%d'), "%Y")

parts <- c("mcr-1", "mcr-2", "mcr-3", "mcr-4", "mcr-5", "mcr-6", "mcr-7", "mcr-8", "mcr-9")

```

## ALDEx2
Comparing relative levels of individual mcr genes between a host/country/year versus all others.
```{r run_ALDEx2}
rerun <- F
if (rerun) {
  print("ALDEx2 for hosts")
  host.res <- full_aldex2(mcr.data, parts, condition_column = 'host')
  print("ALDEx2 for countries")
  country.res <- full_aldex2(mcr.data, parts, condition_column = 'country')
  print("ALDEx2 for years")
  year.res <- full_aldex2(mcr.data, parts, condition_column = 'collection_year')
  print("DONE")
  
  write.csv(host.res, file='data/host_aldex2.csv', row.names=F)
  write.csv(country.res, file='data/country_aldex2.csv', row.names=F)
  write.csv(year.res, file='data/year_aldex2.csv', row.names=F)
} else {
  host.res <- read.csv('data/host_aldex2.csv')
  country.res <- read.csv('data/country_aldex2.csv')
  year.res <- read.csv('data/year_aldex2.csv')
}

```

Let's create some plots.

```{r}
(ga.host <- aldex2_plot(host.res, parts=parts, title='Significant *mcr* levels in hosts', effect_zones = T))
(ga.country <- aldex2_plot(country.res, parts=parts, title='Significant *mcr* levels in countries', effect_zones = T))
(ga.year <- aldex2_plot(year.res, parts=parts, title='Significant *mcr* levels in years', effect_zones = T))

ggsave(plot=ga.host, filename='output/mw_plots/hosts.pdf')
ggsave(plot=ga.country, filename='output/mw_plots/country.pdf')
ggsave(plot=ga.year, filename='output/mw_plots/year.pdf')

```

## PCA
We do PCA on CLR levels.

```{r}
scores <- read.csv('data/mcr_pca_scores_mcr1-9.csv', stringsAsFactors = F, check.names = FALSE)
loadings <- read.csv('data/mcr_pca_loadings_mcr1-9.csv', stringsAsFactors = F, check.names = FALSE)
exp.var <- read.csv('data/mcr_pca_expvar_mcr1-9.csv', stringsAsFactors = F, check.names = FALSE)

colnames(loadings)[1] <- 'PC'
colnames(scores)[6:14] <- paste('PC', 1:9, sep='')

scores$collection_date <- as.Date(scores$collection_date, '%Y-%m-%d')
scores$collection_year <- format(as.Date(scores$collection_date, '%Y-%m-%d'), "%Y")

# add world region to scores df
scores$ADMIN <- str_replace_all(scores$country, c(
  'USA' = 'United States of America',
  'North Macedonia' = 'Macedonia',
  'Tanzania' = 'United Republic of Tanzania',
  'Italia' = 'Italy'
  )
)

scores <- scores %>%
  left_join(countryRegions, by = c("ADMIN" = "ADMIN")) %>%
  mutate(REGION = coalesce(REGION, country))

loadings2 <-loadings %>%
  pivot_longer(parts) %>%
  pivot_wider(names_from=PC, values_from=value)


```

scree plot
```{r}
scree.plot <- ggplot() +
  geom_col(data = exp.var, mapping=aes(`Principal Component`, `Explained Variance`)) +
  ylab('Explained Variance (%)') +
  theme_bw()

ggsave('output/biplots/scree.pdf', plot=scree.plot)
scree.plot
```

biplots

```{r}
axis.labels <- exp.var %>% mutate(
  label = str_c(`Principal Component`, ' (', round(`Explained Variance`, digits=1), '%)'),
) %>% select(label)
rownames(axis.labels) <- exp.var$`Principal Component`

sig.country <- country.res %>%
  filter(we.eBH < 0.05) %>%
  mutate(
    country = str_replace(str_replace(conditions.A.vs.B, ' vs Other', ''), 'Other vs ', '')
  ) %>%
  select(country) %>% distinct()

sig.hosts <- host.res %>%
  filter(we.eBH < 0.05) %>%
  mutate(
    host = str_replace(str_replace(conditions.A.vs.B, ' vs Other', ''), 'Other vs ', '')
  ) %>%
  select(host) %>% distinct()

sig.years <- year.res %>%
  filter(we.eBH < 0.05) %>%
  mutate(
    year = str_replace(str_replace(conditions.A.vs.B, ' vs Other', ''), 'Other vs ', '')
  ) %>%
  select(year) %>% distinct()


```

### simple ones
biplot for significant countries
```{r}
g.pca.country1 <- make_biplot(
  scores = scores, 
  loadings = loadings2, 
  filterby ='country', 
  color='country', 
  x='PC1', y='PC2', 
  sig_val = as.list(sig.country)$country, 
  shape=16, 
  scores_text=T,
  xlabel=axis.labels['PC1', ], ylabel=axis.labels['PC2', ],
  scores_palette = gg_color_hue(n=length(sig.country$country))
  )

ggsave(plot = g.pca.country1, filename = 'output/biplots/biplot_country.pdf', width=15, height=15, units='cm')

g.pca.host1 <- make_biplot(
  scores = scores, 
  loadings = loadings2, 
  filterby ='host', 
  color='host', 
  x='PC1', y='PC2', 
  sig_val = as.list(sig.hosts)$host, 
  shape=16, 
  scores_text=T,
  xlabel=xlabel, ylabel=ylabel,
  scores_palette = gg_color_hue(n=length(as.list(sig.hosts)$host))
  )
ggsave(plot = g.pca.host1, filename = 'output/biplots/biplot_host.pdf', width=15, height=15, units='cm')

g.pca.year1 <- make_biplot(
  scores = scores, 
  loadings = loadings2, 
  filterby ='collection_year', 
  color='collection_year', 
  x='PC1', y='PC2', 
  sig_val = as.list(sig.years)$year, 
  scores_text=T,
  xlabel=xlabel, ylabel = ylabel,
  scores_palette = gg_color_hue(n=length(as.list(sig.years)$year))
  )
ggsave(plot = g.pca.year1, filename = 'output/biplots/biplot_year.pdf', width=15, height=15, units='cm')
```

### Advanced biplots
color by host, shape by year, ellipse for regions
-> but only use those where there are significant groups within category

```{r}
sigs <- list('host'=as.list(sig.hosts)$host, 'collection_year'=as.list(sig.years)$year, 'country'=as.list(sig.country)$country)

g.pca.all <- make_adv_biplot(
  scores=scores, loadings=loadings2, sigs=sigs, color='host', shape='collection_year', ellipses='REGION',
  x='PC1', y='PC2', xlabel = xlabel, ylabel=ylabel, scores_text = T,
  title = 'PCA biplot'
) + labs(tag='d') + theme_update(plot.tag=element_text(face='bold', size=14, color='black'))
ggsave(plot = g.pca.all, filename='output/biplots/biplot_sig_all.pdf', width=30, height=20, units='cm')
ggsave(plot = g.pca.all, filename='output/biplots/biplot_sig_all.png', width=30, height=20, units='cm')
g.pca.all

```

```{r}

g.pca.all <- make_adv_biplot(
  scores=scores, loadings=loadings2, sigs=sigs, color='host', shape='collection_year', ellipses='REGION',
  x='PC1', y='PC2', xlabel = xlabel, ylabel=ylabel, scores_text = F,
  title = 'PCA biplot'
) + theme(
  legend.key.size = unit(1, 'lines')
  ) + labs(tag='d') + theme_update(plot.tag=element_text(face='bold', size=14, color='black'))

ggsave(plot = g.pca.all, filename='output/biplots/biplot_sig_all_legend.pdf', width=30, height=20, units='cm', limitsize = F)
ggsave(plot = g.pca.all, filename='output/biplots/biplot_sig_all_legend.png', width=30, height=20, units='cm', limitsize = F)
```


```{r}
library(ggpubr)
p.mw <-  ggarrange(
  ga.year ,ga.country, ga.host,
  common.legend = T, legend='right',
  nrow = 1, labels='auto'
)
ggsave(plot = p.mw, filename='output/mw_plots/combined.pdf', width = 210, height = 297/3, units = "mm")
ggsave(plot = p.mw, filename='output/mw_plots/combined.png', width = 210, height = 297/3, units = "mm")
p.mw
```

```{r}
ga.pca.combi <- ggarrange(
  g.pca.year1, g.pca.country1, g.pca.host1,
  ncol=1, labels='auto'
)

ggsave(plot = ga.pca.combi, filename='output/biplots/biplot_combined.pdf', height = 210*2, width = 297, units = "mm")
ggsave(plot = ga.pca.combi, filename='output/biplots/biplot_combined.png', height = 210*2, width = 297, units = "mm")

ga.pca.combi
```


```{r}
p.combi <- ggarrange(
  ga.year, g.pca.year1 + ggtitle('Biplot colored by year'),
  ga.country, g.pca.country1 + ggtitle('Biplot colored by country'),
  ga.host, g.pca.host1 + ggtitle('Biplot colored by host'),
  ncol=2, nrow=3, 
  common.legend = T, legend = 'bottom',
  labels='auto'
)
p.combi
ggsave(filename='output/mw_biplot_combined.pdf', plot=p.combi, width=210, height=297, units='mm')
ggsave(filename='output/mw_biplot_combined.png', plot=p.combi, width=210, height=297, units='mm')
```

### PCA without mcr-1 and mcr-9
```{r}
scores2.8 <- read.csv('data/mcr_pca_scores_mcr2-8.csv', stringsAsFactors = F, check.names = FALSE)
loadings2.8 <- read.csv('data/mcr_pca_loadings_mcr2-8.csv', stringsAsFactors = F, check.names = FALSE)
exp.var2.8 <- read.csv('data/mcr_pca_expvar_mcr2-8.csv', stringsAsFactors = F, check.names = FALSE)

colnames(loadings2.8)[1] <- 'PC'
colnames(scores2.8)[6:12] <- paste('PC', 1:7, sep='')

scores2.8$collection_date <- as.Date(scores2.8$collection_date, '%Y-%m-%d')
scores2.8$collection_year <- format(as.Date(scores2.8$collection_date, '%Y-%m-%d'), "%Y")

# add world region to scores df
scores2.8$ADMIN <- str_replace_all(scores2.8$country, c(
  'USA' = 'United States of America',
  'North Macedonia' = 'Macedonia',
  'Tanzania' = 'United Republic of Tanzania',
  'Italia' = 'Italy'
  )
)

scores2.8 <- scores2.8 %>%
  left_join(countryRegions, by = c("ADMIN" = "ADMIN")) %>%
  mutate(REGION = coalesce(REGION, country))

loadings2.8 <-loadings2.8 %>%
  pivot_longer(paste('mcr', 2:8, sep='-')) %>%
  pivot_wider(names_from=PC, values_from=value)

axis.labels2.8 <- exp.var2.8 %>% mutate(
  label = str_c(`Principal Component`, ' (', round(`Explained Variance`, digits=1), '%)'),
) %>% select(label)
rownames(axis.labels2.8) <- exp.var2.8$`Principal Component`


g.pca.year2.8 <- make_biplot(
  scores = scores2.8, 
  loadings = loadings2.8, 
  filterby ='collection_year', 
  color='collection_year', 
  x='PC1', y='PC2', 
  sig_val = as.list(sig.years)$year, 
  shape=16, 
  scores_text=T,
  xlabel=axis.labels2.8['PC1', ], ylabel=axis.labels2.8['PC2', ],
  title='PCA biplot of mcr-2 to mcr-8 colored by year',
  scores_palette = gg_color_hue(n=length(as.list(sig.years)$year))
  )
g.pca.country2.8 <- make_biplot(
  scores = scores2.8, 
  loadings = loadings2.8, 
  filterby ='country', 
  color='country', 
  x='PC1', y='PC2', 
  sig_val = as.list(sig.country)$country, 
  shape=16, 
  scores_text=T,
  xlabel=axis.labels2.8['PC1', ], ylabel=axis.labels2.8['PC2', ],
  title='PCA biplot of mcr-2 to mcr-8 colored by country',
  scores_palette = gg_color_hue(n=length(as.list(sig.country)$country))
  )
g.pca.host2.8 <- make_biplot(
  scores = scores2.8, 
  loadings = loadings2.8, 
  filterby ='host', 
  color='host', 
  x='PC1', y='PC2', 
  sig_val = as.list(sig.hosts)$host, 
  shape=16, 
  scores_text=T,
  xlabel=axis.labels2.8['PC1', ], ylabel=axis.labels2.8['PC2', ],
  title='PCA biplot of mcr-2 to mcr-8 colored by host',
  scores_palette = gg_color_hue(n=length(as.list(sig.hosts)$host))
  )

ggpubr::ggarrange(g.pca.year2.8, g.pca.country2.8, g.pca.host2.8, ncol=1, nrow=3)
ggsave(filename='output/biplots/biplot_combined_mcr-2-8.pdf', height = 210*2, width = 297, units = "mm")
```


# Genomic background
## Load data and create dataframes for plotting
From blast+MEFinder+plasmidFinder
```{r df_MAIN, message=FALSE, warning=FALSE}
df <- read_csv('data/flanks/all_flanks.csv') %>%
  mutate(
    gene_variant = case_when(
      method == 'blastn' ~ str_split(sseqid, '_', simplify = T)[, 1],
      method != 'blastn' ~ sseqid
    ),
    gene = case_when(
      method == 'blastn' ~ str_split(str_split(sseqid, '\\.', simplify = T)[, 1], '_', simplify=T)[, 1],
      method != 'blastn' ~ sseqid
    )
  ) %>%
  separate(
    qseqid, into=c('run_accession', 'contig'), sep=':', remove=F
  ) %>%
  mutate(
    contig_length = as.integer(str_split(contig, '_', simplify = T)[, 4])
    ) %>%
  separate(contig, into=c('node', 'node_number', 'node_other'), sep='_', extra='merge',remove = F) %>%
  mutate(
    contig_number=paste(node, node_number, sep='_'),
    qlen = qend - qstart + 1
  ) %>%
  select(-c(node, node_other))
```

```{r def_window}
window <- 21000
```

```{r df_KMA}
(kmafile=paste0(path, 'data/flanks/all_flank_', window, '_1000.GeneMaskedFlank.fa.phy'))
temp_connection = file(kmafile, 'r')
len = readLines(temp_connection, n=1)
len = as.numeric(len)
close(temp_connection)

kma_data <- as.matrix(read.table(kmafile, fill=T, skip=1, col.names = 1:len, sep='\t', row.names = 1))
mat <- fill_kma.dist(kma_data)

# make tree
kma.clust <- hclust(as.dist(mat), method='ward.D')
tree <- treeio::get.tree(kma.clust)

```

Fetch metadata 
```{r df_METADATA}
update_meta <- F

if(update_meta){
  require(RMariaDB)
  avaDB <- dbConnect(MariaDB(), group='AvA')
  
  query <- "select run_accession, host, country, YEAR(collection_date) as collection_year, refSequence, pct_coverage from get_coverage inner join metadata using(run_accession) where refSequence like 'mcr-%' and pct_coverage>=0.95"
  
  
  metares <- dbSendQuery(avaDB, query)
  metadata <- dbFetch(metares)
  
  dbClearResult(metares)
  dbDisconnect(avaDB)
  unloadNamespace('RMariaDB')
  
  metadata <- metadata %>%
    mutate(
      collection_year = na_if(na_if(collection_year, 0), 1905),
      country = na_if(na_if(country, 'NULL'), 'Not available')
    )

  write_csv(metadata, file = 'data/contigs_metadata.csv')
  } else {
    metadata <- read_csv('data/contigs_metadata.csv')   
  }

```

## Make color palettes
```{r color_palettes}
# function for creating color palettes for categorical groups
make_palette <- function(n, palette=NA){
  
  if (is.na(palette)) {
    hues = seq(15, 375, length = n + 1)
    colors = hcl(h=hues, l=65, c=100)[1:n]
  } else {
    getPalette <- colorRampPalette(brewer.pal(8, palette))
    colors = as.vector(getPalette(n)) 
  }
  return(colors)
}

white.color <- '#FFFFFF'

# mcr gene colors
colors <- make_palette(9)
mcr.gene.colors <- list('NA' = white.color)
for(i in 1:9){
  mcr.gene.colors[[paste('mcr', i, sep='-', collapse='')]] = colors[i]
}

# year coloring scheme
years <- levels(as.factor(unique(metadata$collection_year)))
colors <- make_palette(n=length(years), palette = 'Set3')
year.colors <- list('NA' = white.color)
for(i in 1:length(years)){
  year.colors[[years[i]]] <- colors[i]
}

# country coloring scheme
countries <- unique(metadata$country)
countries <- sort(countries[!is.na(countries)])
colors <- as.list(c(pals::alphabet(), pals::alphabet2()))[1:length(countries)] #make_palette(n=length(countries), palette='Set2')
country.colors <- list('NA' = white.color)
for(i in 1:length(countries)){
  country.colors[[countries[i]]] <- colors[[i]]
}

# host coloring scheme
hosts <- unique(metadata$host)
hosts = sort(hosts[!is.na(hosts)])
colors <- as.list(pals::polychrome(n=length(hosts)))#make_palette(n=length(hosts), palette='Set1')
host.colors <- list('NA' = white.color)
for(i in 1:length(hosts)){
  host.colors[[hosts[i]]] <- colors[[i]]
}

# coloring scheme for plasmids
plasmids <- unique(filter(df, method=='plasmidFinder')$gene)
colors <- make_palette(n=length(plasmids), palette='Dark2')
plasmid.colors <- list('NA' = white.color)
for (i in 1:length(plasmids)) {
  plasmid.colors[[plasmids[i]]] = colors[i]
}

# coloring scheme for mges
mges <- unique(filter(df, method=='MEFinder')$gene)
colors <- make_palette(n=length(mges), palette='Paired')
mge.colors <- list('NA' = white.color)
for (i in 1:length(mges)) {
  mge.colors[[mges[i]]] = colors[i]
}

ele.colors <- c(plasmid.colors, mge.colors)
gen.colors <- c(mcr.gene.colors, plasmid.colors, mge.colors)

```

## Dataframes for plots
Dataframes for plotting trees with metadata
```{r tree.mcr.datas}
tree.mcr.data <- data.frame(tree$tip.label) %>%
  separate(tree.tip.label, into=c('run_accession', 'contig', 'range'), sep=':', remove=F) %>%
  left_join(
    filter(df, method=='blastn'), by=c('run_accession', 'contig')
  ) %>%
  mutate(
    mcr_number = as.factor(as.integer(str_split(gene, '-', simplify = T)[, 2]))
  )

tree.mcr.data2 <- tree.mcr.data %>% left_join(
  distinct(metadata, run_accession, host, country,collection_year), by=c("run_accession"="run_accession")
  )
```

Simplify tree dataframes for genes, years, countries and hosts
```{r treedata}

make_treedata <- function(data, varname, palette, id.vars='tree.tip.label', alpha=1){
  
  tree.data <- reshape2::melt(
      data, id.vars=id.vars, measure.vars=varname, value.name=varname
    )
  tree.data <- tree.data %>% 
    mutate(
      pos=0,
      alpha=alpha
      )
  
  return(tree.data)
}

tree.genes <- make_treedata(tree.mcr.data2, varname='gene', palette=mcr.gene.colors) %>%
  rowwise() %>%
  mutate(color=mcr.gene.colors[gene])

tree.hosts <- make_treedata(tree.mcr.data2, varname='host') %>%
  mutate(
    host = replace_na(host, 'NA')
  ) %>%
  rowwise() %>%
  mutate(color=host.colors[host])

tree.countries <- make_treedata(tree.mcr.data2, varname='country') %>%
  mutate(country = replace_na(country, 'NA')) %>%
  rowwise() %>%
  mutate(color=country.colors[country])

tree.years <- make_treedata(tree.mcr.data2, varname='collection_year') %>%
  mutate(collection_year = as.character(replace_na(collection_year, 'NA'))) %>%
  rowwise() %>%
  mutate(color=year.colors[collection_year])

```


Make dataframe for plotting genomic background
```{r df_GENOMIC_bg}
# create df with mcr genes 'alignment'
sample.df.mcr <- df %>%
  filter(method=='blastn') %>%
  mutate(
    qmid = qlen / 2,
    qstart2 = -qmid,
    qend2 = qmid,
    type = 'mcr'
  )  %>%
  rowwise() %>%
  mutate(
    color =unlist(mcr.gene.colors)[gene], 
    alpha=1)

sample.df.ele <- df[df$window == as.character(window), ] %>%
  filter(method != 'blastn' ) %>%
  mutate(
    qseqid = str_c(run_accession, contig, sep=':')
    #window = as.integer(window)
    ) %>%
  left_join(
    select(sample.df.mcr, qseqid, qstart, qend, qstart2, qend2, qmid, qlen), 
    by='qseqid', 
    suffix=c('.ele', '.mcr')
  ) %>%
  mutate(
    stream = case_when(
      qend.ele < qstart.mcr ~ -1,
      qstart.ele > qend.mcr ~ 1
     ),
    ele_dist = case_when(
      stream == 1 ~ abs(qend.mcr-qstart.ele)- 1,
      stream == -1 ~ abs(qend.ele - qstart.mcr) - 1 
    ),
    qstart2 = case_when(
      stream == 1 ~ qend2 + ele_dist + 1,
      stream == -1 ~ qstart2 - ele_dist - qlen.ele
    ),
    qend2 = qstart2 + qlen.ele + 1,
    type=case_when(
      method == 'plasmidFinder' ~ 'Plasmid',
      method == 'MEFinder' ~ 'MGE'
    )
  ) %>%
  select(-c(qstart.mcr, qend.mcr, qmid, qlen.mcr)) %>%
  rowwise() %>%
  mutate(
    color = unlist(ele.colors)[gene], 
    alpha=1
  )

# contig lengths
sample.df.contigs <- df %>% filter(
  qseqid %in% sample.df.mcr$qseqid
) %>%
  distinct(qseqid, contig_length) %>%
  left_join(sample.df.mcr, by=c('qseqid', 'contig_length')) %>%
  mutate(
    qstart2 = qstart2 - qstart+1,
    qdist = contig_length - qend-1,
    qend2 = qdist + qend2,
    type = 'contig',
    color = '#C0C0C0', # grey,
    alpha = 0.5
  )

df.gene.arrows <- bind_rows(sample.df.contigs, sample.df.ele, sample.df.mcr) %>%
  select(qseqid, gene, qstart2, qend2, type, color, alpha) 

# merge tree tip labels
df.gene.arrows.tree <- data.frame(tree$tip.label) %>%
  separate(tree.tip.label, into=c('run_accession', 'contig', 'range'), sep=':', remove=F) %>%
  mutate(qseqid = str_c(run_accession, contig, sep=':')) %>%
  left_join(df.gene.arrows, by=c('qseqid'))

# df.gene.arrows.tree <- df.gene.arrows.tree[match(df.gene.arrows.tree$tree.tip.label, tree$tip.label), ]

```


## Descriptive plots
What is the length of contigs containing mcr-genes?
```{r mcr_contig_lengths}
max_length <- max(df[df$method=='blastn', 'contig_length'])
step=1000
breaks <- seq(0, max_length+step, by=step)
N_breaks <- length(breaks)
break_labels <- c(
  paste0('<', breaks[2])
)
for(i in 2:N_breaks){
  if (i < N_breaks) {
    break_labels <- c(break_labels, paste(breaks[i], breaks[i+1], sep='-'))
  }
}

g.contig_lengths <- df %>%
  filter(method=='blastn') %>%
  mutate(
    length_bin = cut(as.integer(contig_length), breaks=breaks, labels=break_labels)
  ) %>%
  group_by(gene, length_bin) %>% tally() %>%
  ggplot() +
  geom_col(aes(x=length_bin, y=n, fill=gene)) +
  scale_fill_manual('Gene', values = mcr.gene.colors) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle=90)) +
  ylab('Count') + xlab('Contig length')
g.contig_lengths

ggsave(plot=g.contig_lengths, file='output/contig_plots/contig_lengths.pdf', width=29.7, height=21, units='cm', limitsize = F)

```



What are the size of flanking regions that contains mobile genetic elements or plasmids?
```{r eleCounts_flanks}
mcr.df <- df %>% filter(method == 'blastn') %>% distinct(gene, qseqid)
ele.df <- df %>% filter(method != 'blastn') %>% mutate(qseqid=str_c(run_accession, contig, sep=':'))

count.df <- mcr.df %>% inner_join(ele.df, by='qseqid') %>%
  group_by(gene.x, method, window) %>%
  tally() %>%
  plyr::rename('gene' = 'gene.x', 'Occurence' = 'n')

g.eles <- ggplot(count.df) +
  geom_tile(aes(x=gene, y=as.factor(window), fill=Occurence)) + 
  geom_text(aes(x=gene, y=as.factor(window), label=Occurence),  color='white') + 
  facet_grid(.~method) +
  ylab('maximum flank size') +
  theme_classic()
ggsave('output/contig_plots/flank_ele_sizes.pdf')
```

```{r}
ggpubr::ggarrange(g.contig_lengths, g.eles, ncol = 1, nrow=2, labels='auto')
ggsave('output/contig_plots/contig_flank_ele.pdf', width=25, height=25, units = 'cm')
```



Occurence of plasmids and MGes for each mcr gene
```{r}
g.contig_occurrence <- df[df$window == window, ] %>% filter(method!='blastn') %>%
  mutate(
    qseqid = str_c(run_accession, contig, sep=':'),
    method = str_replace(method, 'MEFinder', 'MobileElementFinder')
  ) %>%
  left_join(df[df$method=='blastn', ], by=c('qseqid', 'run_accession', 'contig', 'node_number')) %>%
  group_by(gene.y, method.x, gene.x) %>%
  tally() %>%
  ggplot() +
  geom_tile(aes(y=gene.y, x=gene.x, fill=n), color='grey') +
  geom_text(aes(y=gene.y, x=gene.x, label = n), color='white') +
  facet_grid(. ~ method.x, scales = 'free_x') + 
  theme_classic() +
  theme(
    axis.text.x = element_text(angle=90),
    ) +
  labs(
    y = expression(paste(italic('mcr'), ' gene')),
    x = 'Plasmid or mobile element'
  )

g.contig_occurrence
ggsave(paste0('output/contig_plots/element_occurence_window', window, '.pdf'), plot=g.contig_occurrence, width=20, height=10, units='cm')
```

```{r}
df[df$window == window, ] %>% filter(method!='blastn') %>%
  mutate(
    qseqid = str_c(run_accession, contig, sep=':'),
    method = str_replace(method, 'MEFinder', 'MobileElementFinder')
  ) %>%
  left_join(df[df$method=='blastn', ], by=c('qseqid', 'run_accession', 'contig', 'node_number')) %>%
  group_by(gene.y, method.x, gene.x) %>%
  tally()
```

## Tree plot 
```{r tree_plot}
gt <- ggtree(tree)#, layout = 'circular')
gt1 <- gt %<+% tree.mcr.data + geom_tippoint(aes(x=x+1, color=gene), size=.75, alpha=.75, shape=15) +
  guides(color=guide_legend(override.aes = aes(size=2)))
gt1
```

## Tree facets
```{r tree_facets_1-4}
p1 <- facet_plot(
  gt,
  panel = 'Gene',
  mapping = aes(x=pos, y=y, fill=gene),
  geom = geom_tile,
  data = tree.genes
) + 
  scale_fill_manual('Gene', values=mcr.gene.colors) + 
  new_scale_fill()

p2 <- facet_plot(
  p = p1,
  panel = 'Year',
  mapping = aes(x=pos, y=y, fill=collection_year),
  geom = geom_tile,
  data = tree.years
) +
  scale_fill_manual('Year', values=year.colors) +
  new_scale_fill()

p3 <- facet_plot(
  p2,
  panel = 'Country',
  mapping = aes(x=pos, y=y, fill=country),
  geom = geom_tile,
  data = tree.countries
) +
  scale_fill_manual('Country', values=country.colors) +
  new_scale_fill()

p4 <- facet_plot(
  p3,
  panel = 'Host',
  mapping = aes(x=pos, y=y, fill=host),
  geom = geom_tile,
  data = tree.hosts
) +
  scale_fill_manual('Host', values=host.colors) +
  theme(legend.position = 'none')

```

Adjust the widths of facets
```{r p4_adjust_widths}

p5 <- ggplot_gtable(ggplot_build(p4))

panels <- c('panel-1-2', 'panel-1-3', 'panel-1-4', 'panel-1-5')
for (panel in panels) {
  p5.col<-p5$layout$l[grep(panel, p5$layout$name)]
  p5$widths[p5.col] = p5$widths[p5.col]*0.075
}

grid::grid.newpage()
grid::grid.draw(p5)

```


## Gene arrows (horizontal lines)
```{r}
# pg <-  ggplot(df.gene.arrows) +
#   geom_segment(
#     aes(x=qstart2, xend=qend2, y=qseqid, yend=qseqid, color=gene), 
#     color=df.gene.arrows$color, 
#     alpha=df.gene.arrows$alpha) + 
#   theme_classic() + 
#   theme(axis.title.x=element_blank(),
#         axis.text.x=element_blank(),
#         axis.ticks.x=element_blank(),
#         axis.title.y=element_blank(),
#         axis.text.y=element_blank(),
#         axis.ticks.y=element_blank()
#         )
# pg

pdf('output/contig_plots/mcr_arrows_individual.pdf', onefile=T)
for(gene in sort(unique(df.gene.arrows.tree[df.gene.arrows.tree$type=='mcr', 'gene']))){

  qseqids <- unique(df.gene.arrows.tree[df.gene.arrows.tree$gene == gene, 'qseqid'])
  qdat <- df.gene.arrows.tree[df.gene.arrows.tree$qseqid %in% qseqids,]
  
  g <- ggplot(
      mapping=aes(x=qstart2, xend=qend2, y=qseqid, yend=qseqid, color=gene)
    ) +
    geom_segment(data=filter(qdat, type=='contig'), color='grey', alpha=.5) +
    geom_segment(data=filter(qdat, type!='contig')) +
    theme_classic() + 
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.title.y=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()
        ) +
    ggtitle(gene)
  
  print(g)
}
dev.off()

```

Add to tree facets
```{r facet_genomic_bg}
LINE_WIDTH=1

p6 <- facet_plot(
  p4 + new_scale_color(),
  panel='Genomic background',
  data = filter(df.gene.arrows.tree, type=='contig'),
  geom = geom_segment,
  mapping = aes(x=qstart2, xend=qend2, y=y, yend=y), 
  color='#C0C0C0',
  alpha = 0.5,
  size = LINE_WIDTH
) + new_scale_color()

p7 <- facet_plot(
  p6,
  panel='Genomic background',
  data = filter(df.gene.arrows.tree, type=='mcr'),
  geom = geom_segment,
  mapping = aes(x=qstart2, xend=qend2, y=y, yend=y, color=gene),
  size = LINE_WIDTH
  ) +
  scale_color_manual('Gene', values=mcr.gene.colors) +
  new_scale_color()

p8 <- facet_plot(
  p7, 
  panel='Genomic background',
  data = filter(df.gene.arrows.tree, type=='Plasmid'),
  geom = geom_segment,
  mapping = aes(x=qstart2, xend=qend2, y=y, yend=y, color=gene),
  size = LINE_WIDTH
  ) +
  scale_color_manual('Plasmid', values=plasmid.colors) +
  new_scale_color()

p9 <- facet_plot(
  p8, 
  panel='Genomic background',
  data = filter(df.gene.arrows.tree, type=='MGE'),
  geom = geom_segment,
  mapping = aes(x=qstart2, xend=qend2, y=y, yend=y, color=gene),
  size = LINE_WIDTH
) +
  scale_color_manual('MGE', values=mge.colors)


# Adjust widths
gp6 <- ggplot_gtable(ggplot_build(p9))

panels <- c('panel-1-2', 'panel-1-3', 'panel-1-4', 'panel-1-5')
for (panel in panels) {
  gp6.col<-gp6$layout$l[grep(panel, gp6$layout$name)]
  gp6$widths[gp6.col] = gp6$widths[gp6.col]*0.08
}

grid::grid.newpage()
grid::grid.draw(gp6)

```

save without legends
```{r}
ggsave(plot = gp6, filename = paste0('output/contig_plots/mcr_tree_window', window, '_noleg.pdf'), width=35, height=21, units = 'cm', limitsize = F)
```


## Legends
```{r MAKE_LEGENDS}

LEGEND_TITLE <- 8
LEGEND_TEXT <- 6
LEGEND_SIZE = 0.5

make_legend <- function(data, x, y, fill, colors, title='', legend_size=0.5, legend_text=6, legend_title=8, byrow=F, nrow=NULL, ncol=NULL, legend.direction='vertical', legend.box='vertical', title.position='top'){
  
  g.leg <- ggplot(data=data, mapping=aes_string(x=x, y=y, fill=fill)) +
    geom_tile() + 
    scale_fill_manual(title, values=colors) +
    guides(
      fill=guide_legend(
        override.aes = list(size=legend_size),
        byrow=byrow,
        nrow=nrow,
        ncol=ncol,
        title.position=title.position
      )
    ) + 
    theme(
      legend.text = element_text(size=legend_text),
      legend.title = element_text(size=legend_title),
      legend.direction = legend.direction, 
      legend.box = legend.box,
      axis.text.y = element_blank()
    )
  
  leg <- get_legend(g.leg)
  return(leg)
}

mcr.legend <- make_legend(data=tree.genes, colors=mcr.gene.colors, x='pos', y='tree.tip.label', fill='gene', title = 'Gene')
plot_grid(mcr.legend)

plasmid.legend <- make_legend(data=filter(sample.df.ele, type=='Plasmid'), colors=plasmid.colors, x='qseqid', y='window', fill='sseqid', title='Plasmid')
plot_grid(plasmid.legend)

mge.legend <- make_legend(data=filter(sample.df.ele, type=='MGE'), colors=mge.colors, x='qseqid', y='window', fill='sseqid', title='MGE') 
plot_grid(mge.legend)

(ele.legend1 <- plot_grid(mcr.legend, NULL, plasmid.legend, ncol=1, align='hv', rel_heights = c(1, -0.1, 1)))
(ele.legend2 <- plot_grid(ele.legend1, NULL, mge.legend, ncol=3, rel_widths=c(1, -.5, 1)))

n_row = 5
year.legend <- make_legend(data=tree.years, x='pos', y='tree.tip.label', fill='collection_year', title='Year', colors=year.colors, nrow=n_row, legend.direction='horizontal', byrow=T)
plot_grid(year.legend)

country.legend <- make_legend(data=tree.countries, x='pos', y='tree.tip.label', fill='country', colors=country.colors, title='Country', nrow=n_row)
plot_grid(country.legend)

host.legend <- make_legend(data=tree.hosts, x='pos', y='tree.tip.label', fill='host', colors=host.colors, title='Host', nrow=n_row)
plot_grid(host.legend)

meta.legend <- plot_grid(year.legend, NULL, country.legend, NULL, host.legend, nrow=1, ncol=5, rel_widths=c(1, -.75, 1, -.5, 1), align='hv')
```

with legends
```{r}
par(xpd=NA, bg='transparent')
gp7 <- plot_grid(gp6, NULL, ele.legend2, ncol=3, rel_widths=c(2, -.2, 1))
gp8 <- plot_grid(gp7, NULL, meta.legend, nrow=3, rel_heights=c(2, -.2, 1))
# cowplot::plot_grid(gp6, legend.genes, leg1, leg2, nrow=4, rel_heights = c(1, .25, .45, .75))
ggsave(paste0('output/contig_plots/mcr_tree_window', window, '.pdf'), plot=gp8, width=35, height=21, units = 'cm', limitsize = F)
ggsave(paste0('output/contig_plots/mcr_tree_window', window, '.png'), width=35, height=21, units = 'cm', limitsize = F)
dev.off()
```
